---
title: "AddPointCountSites"
author: "Patrick D. lorch"
date: "2025-04-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Boundary import



```{r bounds}

```

## GRTS

### MGRS grid cell selection

To follow RMH Handbook, starting on pg 54, we download an MGRS grd from https://mgrs-data.org/ (.gdb version) or
https://earth-info.nga.mil/index.php?dir=coordsys&action=mgrs-1km-polyline-dloads (.shp version)

UTM 11N (11S?) covers our part of CA.

That is stored in SharePoint at "~Southern Sierra Research Station - Documents\Projects\TNC Randall Preserve Projects\Maps & GIS\mgrs_11s_100m" and 
""~Southern Sierra Research Station - Documents\Projects\TNC Randall Preserve Projects\Maps & GIS\MGRS_1km_11S_unprojected"

To make the leaflet map work currently, 

* you run it after you find the UTM id for the 1km block(s) including your sites. 
  * For Randall ranches this is LV and LU
* In the leaflet map, you can hover over the cells that overlap your sites and see the MGRS ID for the cell

**Note:** MGRS desciption: 11SLV3713, where first 3 are UTM grid ID, middle are MGRS 1km grid ID, and last four are CCRR where CC is column ID (increasing left to right) and RR is Row id (increasing bottom to top)

The first Leaflet map is just for figuring out what MGRS IDs you need to include.

The plots are generally just to help check you are getting what you expect.


```{r mgrspolys}
library(dplyr)
library(leaflet)
library(leaflet.extras2)
library(sf)
library(geojsonsf)
library(plotrix)
library(htmlwidgets)

# Things you need to set for each run
ncells = 4 # number of cells to divide MGRS cells by for GRTS sampling
MGRS_1km_11S_unprojected = 
  st_read(dsn = file.path("../..",
                    "Maps & GIS",
                    "MGRS_1km_11S_unprojected"),
          layer = "MGRS_1km_11S_unprojected")
# Subset to one or more km^2 block for easier plotting
MGRS_1km_11S_subset = 
  MGRS_1km_11S_unprojected %>%
  filter(kmSQ_ID %in% c("LV", "LU"))
center = st_coordinates(st_centroid(summarize(MGRS_1km_11S_subset)))
zoom = 10

```

#### Sampling larger areas

If a single polygon you are sampling is large enough to contain a decent GRTS sample, this method will work.

```{r mgrsbigpolys}
# For sampling the ranches

ranch_polygons =
  st_read(dsn = file.path("../..",
                          "Maps & GIS",
                          "2023 (complete GIS files)",
                          "RanchPolygons"),
          layer = "Polygons")
ranch_polygons =
  st_zm(ranch_polygons, drop = T, what = "ZM")
# get in same coordinate system
ranch_polygons = st_transform(ranch_polygons, st_crs(MGRS_1km_11S_subset))

pal <- colorFactor(
  palette = "magma",
  domain = ranch_polygons$Name,
  na.color = "#737373")

m = leaflet() %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles('Esri.WorldImagery', 
                   group = "Esri World Imagery") %>%
  setView(center[1], center[2], zoom) %>%
  addPolygons(data = MGRS_1km_11S_subset,
              fill = F,
              color="black", 
              opacity = 0.8,
              label = ~MGRS,
              group = "MGRS") %>%
  addPolygons(data = ranch_polygons,
              fill = F,
              color= ~pal(Name),
              label = ~Name,
              opacity = 0.8,
              group = "Ranch") %>%
  addLayersControl(
    baseGroups = c("Esri World Imagery", "OSM (default)"),
    overlayGroups = c("MGRS", "Ranch"),
    position = "bottomright",
    options = layersControlOptions(collapsed = TRUE)
    )

m
saveWidget(m, file = "MGRSgridcellpickerBig.html")
```

#### Sampling smaller areas

If a single polygon you are sampling is NOT large enough to contain a decent GRTS sample, this method will work better. We use this for adding GRTS points "in and around" the spot mapping plots.

```{r mgrssmallpolys}
subset_polys = c("Tollhouse Ranch")
focal_ranch_polygon = filter(ranch_polygons,
                             Name %in% subset_polys)
focal_ranch_polygon_6340 = st_transform(focal_ranch_polygon, 6340)
center = st_coordinates(st_centroid(summarize(focal_ranch_polygon)))
zoom = 12

# Polygons of interest for narrowing down GRTS point sample frame
##  First get spotmapping plots into one sfdf
focal_poly_name1 = "Spotmapping2025"
focal_polygons1 =
  st_read(dsn = file.path("../..",
                          "Maps & GIS",
                          "SpotMappingShapefileExport",
                          "Spot_mapping_plots_added_in_2025"),
          layer = "Spot_mapping_plots_added_in_2025") 
# Need to remove Z dimension which this shapefile was saved with
focal_polygons1 = st_zm(focal_polygons1, drop = T, what = "ZM")
# get in same coordinate system
focal_polygons1 = st_transform(focal_polygons1, 
                              st_crs(MGRS_1km_11S_subset))

focal_poly_name2 = "SpotmappingPre2025"
focal_polygons2 = 
  st_read(dsn = file.path("../..",
                          "Maps & GIS",
                          "SpotMappingShapefileExport",
                          "TNC_RP_2024_WFL1_spotmappingPre2025"),
          layer = "Spot_Mapping_Plots")
# get in same coordinate system
focal_polygons2 = st_transform(focal_polygons2, 
                              st_crs(MGRS_1km_11S_subset))
focal_polygons = bind_rows(focal_polygons1, focal_polygons2)

# Rename duplicate created by casting multipolygon to polygon
focal_polygons$PlotName[which(focal_polygons$PlotName == "BR-A")[2]] =
  "BR-A2"
focal_polygons = focal_polygons %>%
  mutate(YearAdded = c(rep(2025, 2), 
                       rep(2024, nrow(focal_polygons2)))) %>%
  filter(!(PlotName %in% c("Exclusion", "Exclusion2")))
## Subset this by ranch(s)
focal_polygons = 
  st_intersection(focal_polygons,
               focal_ranch_polygon)


pal <- colorFactor(
  palette = "magma",
  domain = focal_polygons$PlotName,
  na.color = "#737373")

m = leaflet() %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles('Esri.WorldImagery', 
                   group = "Esri World Imagery") %>%
  setView(center[1], center[2], zoom) %>%
  addPolygons(data = MGRS_1km_11S_subset,
              fill = F,
              color="black", 
              opacity = 0.8,
              label = ~MGRS,
              group = "MGRS") %>%
  addPolygons(data = focal_polygons,
              fill = F,
              color = ~pal(PlotName),
              label = ~PlotName,
              opacity = 0.8,
              group = "Spotplot") %>%
  addPolygons(data = focal_ranch_polygon,
              fill = F,
              color = "purple",
              opacity = 0.75,
              group = "Ranch") %>%
  addLayersControl(
    baseGroups = c("Esri World Imagery", "OSM (default)"),
    overlayGroups = c("MGRS", "Spotplot", "Ranch"),
    position = "bottomright",
    options = layersControlOptions(collapsed = TRUE)
    )

m
saveWidget(m, file = "MGRSgridcellpickerSmall.html")

```


### Vector grid 

Develop a grid to put a certain density of points within each 1 km MGRS cell and eliminate points near and outside the boundary.

Cannot use meters with negative buffers in sf if use_s2 is True so using estimating fractional degrees. This number can be changed if you want to remove more boundary points based on plots.

We probably need to increase small_poly_buffer_m until we have around 40 potential GRTS points per ranch.

```{r vectorgrid}
library(sf)

# small_poly_buffer_m = 0
small_poly_buffer_m = 700
# neg_buffer_deg = (-0.001)
neg_buffer_m = (-100)

focal_poly = focal_polygons
# Narrow down area grid is built for using st_intersection
# focal_poly = focal_polygons %>% filter(Name == focal_poly_name)
MGRS_1km_11S_intersect = st_intersection(MGRS_1km_11S_subset, st_buffer(focal_poly, small_poly_buffer_m))
# Get list of MGRS blocks of interest
MGRS_1km_11S_intersect_ids = unique(MGRS_1km_11S_intersect$MGRS)
# Subset the MGRS grid
MGRS_1km_11S_subset3 = MGRS_1km_11S_subset %>%
  filter(MGRS %in% MGRS_1km_11S_intersect_ids)
plot(MGRS_1km_11S_intersect["MGRS"])
# plot(MGRS_1km_11S_intersect["MGRS"])

plot(st_as_sfc(st_bbox(MGRS_1km_11S_subset3)), axes = T)
plot(MGRS_1km_11S_subset3["MGRS"], axes = T, add = T)

MGRS_1km_11S_subset3_6340 = st_transform(MGRS_1km_11S_subset3, crs = 6340)

# Start here to redo debugging plot

plot(st_as_sfc(st_bbox(MGRS_1km_11S_subset3_6340)), axes = T)
plot(MGRS_1km_11S_subset3_6340["MGRS"], add = T)
# Find width and length of MGRS grid cells
MGRS_height = 
  diff(as.numeric(range(substr(MGRS_1km_11S_subset3_6340$MGRS, 8, 9)))) + 1
MGRS_width = 
  diff(as.numeric(range(substr(MGRS_1km_11S_subset3_6340$MGRS, 6, 7)))) + 1

MGRS_1km_11S_subset3_grid = 
  st_make_grid(MGRS_1km_11S_subset3_6340, 
               # cellsize = ncells *c(MGRS_height, MGRS_width),
               n = ncells *c(MGRS_height, MGRS_width),
               what = "centers")
plot(MGRS_1km_11S_subset3_grid, add = T)

# Remove points outside of buffered focal poly and within ~100m inside ranch boundary
focal_poly_6340 = focal_poly %>%
  st_buffer(small_poly_buffer_m) %>%
  st_transform(6340)
plot(focal_poly_6340, add = T, col = NA, lwd = 2)
MGRS_1km_11S_subset4_grid = 
  st_intersection(x = MGRS_1km_11S_subset3_grid,
                  y = focal_poly_6340)

plot(MGRS_1km_11S_subset4_grid, pch = 19, add = T)
# MGRS_1km_11S_subset4_grid = 
#   st_intersection(x = MGRS_1km_11S_subset3_grid,
#                   y = focal_poly_6340)
# Get points within < 100m of ranch boundary
MGRS_1km_11S_subset5_grid =
  st_intersection(x = MGRS_1km_11S_subset4_grid,
                  y = st_buffer(focal_ranch_polygon_6340,
                            dist = neg_buffer_m))
plot(st_buffer(focal_ranch_polygon_6340,
               dist = neg_buffer_m), 
     col = NA,
     add = T)

plot(MGRS_1km_11S_subset5_grid, pch = 19, col = "red", add = T)
# Check how many potential points
length(MGRS_1km_11S_subset5_grid)

# Need Leaflet map to find and eliminate trees withing 50m
center = st_coordinates(st_centroid(summarize(MGRS_1km_11S_subset3)))
zoom = 12
# starting_points = MGRS_1km_11S_subset4_grid
starting_points = st_transform(MGRS_1km_11S_subset5_grid, 4326)

m2 = leaflet() %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles('Esri.WorldImagery', 
                   group = "Esri World Imagery") %>%
  setView(center[1], center[2], zoom) %>%
  addPolygons(data = MGRS_1km_11S_subset3,
              fill = F,
              color = "black", 
              opacity = 0.8,
              label = ~MGRS,
              group = "MGRS") %>%
  addPolygons(data = focal_polygons,
              fill = F,
              color= ~pal(PlotName), 
              label = ~PlotName,
              opacity = 0.8,
              group = "Spotplot") %>%
              # color= ~pal(Name), 
              # label = ~Name,
              # group = "Ranch") %>%
  addPolygons(data = focal_ranch_polygon,
              fill = F,
              color = "red",
              opacity = 0.8,
              label = ~Name,
              group = "Ranch") %>%
  addCircles(data = starting_points,
                   radius = 50,
                   fill = F,
                   label = 1:length(starting_points),
                   group = "Points") %>%
  addLayersControl(
    baseGroups = c("Esri World Imagery", "OSM (default)"),
    overlayGroups = c("MGRS", "Spotplot"),
    # overlayGroups = c("MGRS", "Ranch"),
    position = "bottomright",
    options = layersControlOptions(collapsed = TRUE)
  # addLayersControl(
  #   baseGroups = c("Esri World Imagery", "OSM (default)"),
  #   overlayGroups = c("MGRS", "Ranch", "Points"),
  #   position = "bottomright",
  #   options = layersControlOptions(collapsed = TRUE)
    ) %>%
  addScaleBar(position = "bottomleft")


m2
saveWidget(m, file = "MGRSgridcelleliminator.html")
# Ordered list of points within 50 m of trees
tree_points = c(5,9,11,13:16,18,19, 28, 36, 40,44)
ending_points = st_transform(starting_points[-tree_points], 6340)
plot(ending_points, pch = 19, col = "green", add = T)
sampling_points = ending_points

```

### Grts drop

The code from p. 56 of the RMH Handbook is out of date and useless.

```{r grts}
library(spsurvey)
library(sf)

# Set values for sample density here
main_points = 15
over_points = 10
point_ID_prefix = subset_poly

grtsRanchX = list(None=list(panel=c(Panel1 = main_points),
                           seltype = "Equal", 
                           over = over_points))

focal_grts = grts(st_as_sf(sampling_points), 
                  n_base = 15,
                  mindis = 250,
                  n_over = 10,
                  DesignID = point_ID_prefix, # Could be/add MGRS ID
                  SiteBegin = 1)
plot(focal_grts$sites_base$x, pch = 19, col = "blue", add =T)
plot(focal_grts$sites_over$x, pch = 19, col = "orange", add =T)

```

