---
title: "PointCountMonitoring"
author: "Patrick D. lorch"
date: "2022-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data

Set up for each year's analysis.

```{r data}
library(readxl)
library(tidyr)
library(dplyr)
library(lubridate)
```

### 2021

Annie worked on the 2021 TNC Caliente bird survey data file to clean up mistakes in the Species codes.  Other changes includedThose done in R below.

```{r data21}

caliente_location_codes = data.frame(site_code = c("THO","THG","CBG","CO","CC","THR","BR"),
                                        site_name = c("Tollhouse – Oak Savanna",
                                                      "Tollhouse – Grassland",
                                                      "Caliente – Buckwheat/Goldenbush Scrub",
                                                      "Caliente – Oak Woodland",
                                                      "Caliente Creek - Riparian",
                                                      "Walker Basin Creek (Tollhouse) – Riparian",
                                                      "Beard Ranch - Riparian"),
                                        id = 1:7)
write.csv(caliente_location_codes, "caliente_location_codes.csv")

surveyor = data.frame(initials = c("MW", "EJ", "KS", "NJ", "RV", "MP", 
                                   "SR", "RET", "PL", "AJR", "ALM", 
                                   "AJP", "BH", "SD", "RJ", "DT", "LR",
                                   "FC"),
                      name = c("Mary Whitfield",
                               "Edwin Jacobo",
                               "Kristie Stein",
                               "Nidia Jaime",
                               "Rachel Voss",
                               "Maggie Pasanen",
                               "Sean Rowe",
                               "Reed Tollefson",
                               "Patrick Lorch",
                               "Alexander Robinson",
                               "Annie Meyer",
                               "Aidan Place",
                               "Brian Hofstetter",
                               "Sara Dart",
                               "RJ Baltierra",
                               "Danny Tipton",
                               "Lauren Roux",
                               "Fen Conway"),
                      id = 1:18)
write.csv(surveyor, "surveyor.csv")

# Read in Point count data
tnc_pc_2021 <- read_excel("C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/TNC Bird Surveys - Caliente_WB/Data/2021/ALM_Mod_TNC Point Count Data 2021_110521.xlsx",sheet = "Sheet3", col_types = c("numeric","text", "text", "date", "text", "text","numeric", "numeric", "numeric","numeric", "numeric", "numeric","numeric", "numeric", "numeric","text", "text", "text", "text", "text","text", "text", "numeric", "numeric","text", "text", "text", "text"))

# Sanity check
table(tnc_pc_2021$Species, tnc_pc_2021$Location_Code)
# Check for any unusual codes like UEFL
which(tnc_pc_2021_all$Species == "UEFL")
# Check distance
hist(tnc_pc_2021$Distance)

# Get data for birds not at points
tnc_88_2021 <- read_excel("C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/TNC Bird Surveys - Caliente_WB/Data/2021/ALM_Mod_TNC Point Count Data 2021_110521.xlsx",sheet = "88 Species",
                          col_types = c("date","text", "text", "text", "text", 
                                        "text", "text", "text", "numeric", 
                                        "text", "numeric", "numeric", "text",
                                        "text"))
# Rename column from this table to allow habitat assignment
names(tnc_88_2021)[2] = "Location_Code"
# Combine for some analyses but add a column that tracks where detected
tnc_combined_2021 = tnc_pc_2021 %>%
  bind_rows(tnc_88_2021, .id = "point_vs_88")

# Set dummy variable for No birds at a point (NOBI)
tnc_combined_2021 = tnc_combined_2021 %>%
  mutate(nobi = ifelse(Species == "NOBI", 1, 0))

# Mark unknowns
tnc_combined_2021 = tnc_combined_2021 %>%
  mutate(known = substr(Species,1,1) != "U")

# Add habitat type variable

tnc_combined_2021= tnc_combined_2021 %>%
  mutate(Habitat = case_when(Location_Code == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             Location_Code == "CO" ~ "Oak Woodland",
                             Location_Code == "THG" ~ "Grassland",
                             Location_Code == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian"))
tnc_combined_2021 = tnc_combined_2021 %>%
  mutate(Season = case_when(Visit == "3" ~ "Summer",
                            Visit == 1 | Visit == 2 ~ "Spring",
                            TRUE ~ "Fall"))
           

# summaries 
sp_by_hab_21 = tnc_combined_2021 %>%
  filter(known == T, nobi == 0) %>%
  group_by(Species, Habitat) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_21, "sp_by_hab21.csv")

sp_counts_by_hab_21 = colSums(sp_by_hab_21 != 0, na.rm = T)
write.csv(as.data.frame(sp_counts_by_hab_21), "sp_counts_by_hab21.csv")


sp_by_hab_season = tnc_combined_2021 %>%
  filter(known == T, nobi == 0) %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_season, "sp_by_hab_season.csv")

# Removing THG-09 which is not really a grassland
# Requires eliminating two records from the 88 section
#   These were identified with the identifier below
# *** This needs to be done anyplace else you want to have a strict definition
#     of Grassland. ***
tnc_combined_woTHG9_2021 = tnc_combined_2021 %>%
  filter(known == T, nobi == 0, !(Point_Name %in% c("THG-09",
                                        "THG-10",
                                        "THG-11",
                                        "THG-12",
                                        "THG-13",
                                        "THG-14",
                                        "THG-15",
                                        "THG-16",
                                        "THG-17",
                                        "THG-18")),
                                        !(point_vs_88 == 2 &
                                                Location_Code == "THG" &
                                                Observer == "KS" &
                                                Date == "2021-04-24"))

sp_by_hab_season_woTHG9 = tnc_combined_woTHG9_2021 %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_season_woTHG9, "sp_by_hab_season_woTHG9.csv")

sp_by_hab_season_wo88 = tnc_combined_2021 %>%
  filter(known == T, nobi == 0, point_vs_88 == 1) %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_season_wo88, "sp_by_hab_season_wo88.csv")

# By using tnc_pc_2021 we eliminate 88s and include surveys where they did not 
# see any birds (NOBI)
caliente_observer_2021 = tnc_pc_2021 %>%
  select("Date2", "Location_Code", "Observer") %>%
  unique() %>%
  left_join(caliente_location_codes, by = c("Location_Code" = "site_code")) %>%
  left_join(surveyor, by = c("Observer" = "initials")) %>%
  select(Date = Date2, Site = site_name, Surveyor = name)
write.csv(caliente_observer_2021, "caliente_observer_2021.csv")

bird_sp = read.csv("birds_202210041059.csv")
bird_sp = bird_sp %>%
  select(spec, commonname, sciname)

sp_list_points = tnc_combined_2021 %>%
  filter(known == T, nobi == 0, point_vs_88 == 1) %>%
  select(Species, Season)%>%
  distinct() %>%
  arrange(Season, Species) %>%
  left_join(bird_sp, by = c("Species" = "spec"))

sp_list_points_by_season = sp_list_points %>%
  mutate(count = 1) %>%
  pivot_wider(names_from = Season, values_from = count)
# generate totals by season if time permits
write.csv(sp_list_points_by_season, "sp_list_points_by_season.csv")

sp_list_88 = tnc_combined_2021 %>%
  filter(known == T, nobi == 0, point_vs_88 == 2) %>%
  select(Species, Season)%>%
  distinct() %>%
  arrange(Season, Species) %>%
  left_join(bird_sp, by = c("Species" = "spec"))

sp_list_88_by_season = sp_list_88 %>%
  mutate(count = 1) %>%
  pivot_wider(names_from = Season, values_from = count)
write.csv(sp_list_88_by_season, "sp_list_88_by_season.csv")

# Anti join to find unique species in 88s
sp_list_88_21_unique = sp_list_88 %>%
  anti_join(sp_list_points, by = "Species") %>%
  mutate(count = 1) %>%
  pivot_wider(names_from = Season, values_from = count)
write.csv(sp_list_88_21_unique, "sp_list_88_21_unique.csv")


# Not sure what to do with this.  It came from an example I found someplace.
library(mefa4)
library(vegan)


count2021 = Xtab(Number ~ Habitat + Species +Season, tnc_pc_2021)
# count2021 = as.matrix(count2021)
count2021Fall = specnumber(count2021$Fall, MARGIN = 1)
count2021Spring = specnumber(count2021$Spring, MARGIN = 1)
count2021Summer = specnumber(count2021$Summer, MARGIN = 1)


```

### 2022

Most changes were done by Sasha in spreadsheets.

#### Interim reports for 2022

```{r data22}

# Read in Point count data
tnc_pc_2022 <- read_excel("C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/TNC Bird Surveys - Caliente_WB/Data/2022/TNC22_Sasha_wbook.xlsx", 
sheet = "21-22w88", col_types = c("numeric", 
    "text", "text", "date", "text", "text", 
    "numeric", "numeric", "numeric", 
    "numeric", "numeric", "numeric", 
    "numeric", "numeric", "numeric", 
    "text", "numeric", "text", "text", 
    "text", "text", "text", "text", "numeric", 
    "numeric", "text", "numeric", "text", 
    "text", "text"))

# Sanity check
table(tnc_pc_2022$Species, tnc_pc_2022$Location_Code)

# Set dummy variable for No birds at a point (NOBI)
tnc_combined_2022 = tnc_pc_2022 %>%
  mutate(nobi = ifelse(Species == "NOBI", 1, 0))

# Mark unknowns
tnc_combined_2022 = tnc_combined_2022 %>%
  mutate(known = substr(Species,1,1) != "U")

# Get just 2022 data
tnc_combined_2022 = tnc_combined_2022 %>%
  filter(Date > "2021-12-31")

# Add habitat type variable

tnc_combined_2022= tnc_combined_2022 %>%
  mutate(Habitat = case_when(Location_Code == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             Location_Code == "CO" ~ "Oak Woodland",
                             Location_Code == "THG" ~ "Grassland",
                             Location_Code == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian"))
tnc_combined_2022 = tnc_combined_2022 %>%
  mutate(Season = case_when(Visit == "3" ~ "Summer",
                            Visit == 1 | Visit == 2 ~ "Spring",
                            TRUE ~ "Fall"),
         point_vs_88 = case_when(Point_Name == "88" ~ 2,
                                 TRUE ~ 1)
         )
           

# summaries 
sp_by_hab_22 = tnc_combined_2022 %>%
  filter(known == T, nobi == 0) %>%
  group_by(Species, Habitat) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_22, "sp_by_hab_22.csv")

sp_counts_by_hab_22 = colSums(sp_by_hab_22 != 0, na.rm = T)
write.csv(as.data.frame(sp_counts_by_hab_22), "sp_counts_by_hab_22.csv")

sp_by_hab_season_22 = tnc_combined_2022 %>%
  filter(known == T, nobi == 0) %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_season_22, "sp_by_hab_season_22.csv")

sp_by_hab_season_wo88_22 = tnc_combined_2022 %>%
  filter(known == T, nobi == 0, point_vs_88 == 1) %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_season_wo88_22, "sp_by_hab_season_wo88_22.csv")

caliente_observer_2022 = tnc_combined_2022 %>%
  select("Date", "Location_Code", "Observer") %>%
  unique() %>%
  left_join(caliente_location_codes, by = c("Location_Code" = "site_code")) %>%
  left_join(surveyor, by = c("Observer" = "initials")) %>%
  select(Date = Date, Site = site_name, Surveyor = name)
write.csv(caliente_observer_2022, "caliente_observer_2022.csv")

bird_sp = read.csv("birds_202210041059.csv")
bird_sp = bird_sp %>%
  select(spec, commonname, sciname)

sp_list_points_22 = tnc_combined_2022 %>%
  filter(known == T, nobi == 0, point_vs_88 == 1) %>%
  select(Species, Season)%>%
  distinct() %>%
  arrange(Season, Species) %>%
  left_join(bird_sp, by = c("Species" = "spec"))

sp_list_points_22_by_season = sp_list_points_22 %>%
  mutate(count = 1, nobi == 0) %>%
  pivot_wider(names_from = Season, values_from = count)
# generate totals by season if time permits
write.csv(sp_list_points_22_by_season, "sp_list_points_22_by_season.csv")

sp_list_88_22 = tnc_combined_2022 %>%
  filter(known == T, nobi == 0, point_vs_88 == 2) %>%
  select(Species, Season)%>%
  distinct() %>%
  arrange(Season, Species) %>%
  left_join(bird_sp, by = c("Species" = "spec"))

sp_list_88_22_by_season = sp_list_88_22 %>%
  mutate(count = 1) %>%
  pivot_wider(names_from = Season, values_from = count)
write.csv(sp_list_88_22_by_season, "sp_list_88_22_by_season.csv")

# Anti join to find unique species in 88s
sp_list_88_22_unique = sp_list_88_22 %>%
  anti_join(sp_list_points_22, by = "Species") %>%
  mutate(count = 1) %>%
  pivot_wider(names_from = Season, values_from = count)
write.csv(sp_list_88_21_unique, "sp_list_88_21_unique.csv")

```


#### Community analyses

We want to compare habitat types in a dendrogram.

Following the vegan community vignette:
https://www.mooreecology.com/uploads/2/4/2/1/24213970/vegantutor.pdf

```{r community22}
#install.packages("vegan")
library(MASS)
library(vegan)
library(tidyr)
library(dplyr)
library(plotrix)
library(indicspecies)
library(ggplot2)

tnc_site_by_species_all = tnc_combined_2021 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101) %>%
  group_by(Species, Point_Name) %>%
  count() %>%
  pivot_wider(names_from = Species, values_from = n)

tnc_site_by_species_all[is.na(tnc_site_by_species_all)] = 0
site_species_all=data.frame(tnc_site_by_species_all[,-1])

tnc_site_by_species = tnc_site_by_species_all %>%
  filter(!(Point_Name %in% c("THG-09",
                                        "THG-10",
                                        "THG-11",
                                        "THG-12",
                                        "THG-13",
                                        "THG-14",
                                        "THG-15",
                                        "THG-16",
                                        "THG-17",
                                        "THG-18")))

tnc_site_by_species[is.na(tnc_site_by_species)] = 0
site_species=data.frame(tnc_site_by_species[,-1])

sites_2021 = tnc_site_by_species %>% pull(Point_Name)
transects_2021 = sapply(strsplit(sites_2021,"-"), `[`, 1)
habitat_2021 = as.data.frame(transects_2021) %>%
  mutate(Habitat = case_when(transects_2021 == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             transects_2021 == "CO" ~ "Oak Woodland",
                             transects_2021 == "THG" ~ "Grassland",
                             transects_2021 == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian")) %>%
  dplyr::select(Habitat)
col_21_23 = habitat_2021 %>%
  mutate(color = case_when(habitat_2021 == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           habitat_2021 == "Oak Woodland" ~ "grey",
                           habitat_2021 == "Grassland" ~ "green",
                           habitat_2021 == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue")) %>%
    dplyr::select(color)

tnc_2021_shannon = diversity(site_species)

tnc_2021.dis = vegdist(site_species)
tnc_2021.mds0 = isoMDS(tnc_2021.dis)
stressplot(tnc_2021.mds0, tnc_2021.dis)
# An Ordination plot
p1 = ordiplot(tnc_2021.mds0, type = "points")
row.names(p1$sites) = sites_2021
text(p1, "sites", col = "red")

# Ordination using NMDS and random starting points
tnc_2021.mds = metaMDS(site_species, trace = F)
tnc_2021.mds
row.names(tnc_2021.mds$points) = sites_2021
p1 = plot(tnc_2021.mds, disp = "sp")
# This should reduce business of species labeling, but it takes forever with this ordiplot
# identify(p1, "sp")
hab_fit_2021 = envfit(tnc_2021.mds, habitat_2021, permu = 999)
p1 = plot(tnc_2021.mds, disp = "sp")
plot(hab_fit_2021, p.max = 0.1)

# Some Ordination plots
ordiplot(tnc_2021.mds, type = "n")
ordihull(tnc_2021.mds, groups = habitat_2021$Habitat, draw="polygon",
         col = c("yellow", "green", "brown", "grey", "blue"), label=F)
ordiellipse(tnc_2021.mds, groups = habitat_2021$Habitat, draw="polygon",
         col = c("yellow", "green", "brown", "grey", "blue"), label=F)
# ordispider(tnc_2021.mds, groups=habitat_2021$Habitat)
orditorp(tnc_2021.mds, display="species", col="red", air=0.01)
orditorp(tnc_2021.mds, display="sites", col = col_2021$color,
   air=0.01,cex=1.25)


tnc_2021.clus = hclust(tnc_2021.dis, "single")
plot(tnc_2021.clus, labels = sites_2021)

tnc_2021.cluc = hclust(tnc_2021.dis, "complet")
plot(tnc_2021.cluc, labels = sites_2021)

tnc_2021.clua = hclust(tnc_2021.dis, "average")
plot(tnc_2021.clua, labels = sites_2021)

range(tnc_2021.dis)

cor(tnc_2021.dis, cophenetic(tnc_2021.clus))
cor(tnc_2021.dis, cophenetic(tnc_2021.cluc))
cor(tnc_2021.dis, cophenetic(tnc_2021.clua))

# Following my methods from CMP point count talk
tnc_2021.euc = vegdist(site_species, method = "euc")
tnc_2021.ward = hclust(tnc_2021.euc, method = "ward.D")
plot(tnc_2021.ward, labels = sites_2021)
cor(tnc_2021.euc, cophenetic(tnc_2021.ward))

indval_2021 = multipatt(site_species, habitat_2021$Habitat, control = how(nperm = 999))
summary(indval_2021)


## By season
# Need to eliminate THG-09 - 18 for spring 2021 and possibly 2022

## Spring
tnc_site_by_species_spring = tnc_combined_2021 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101,
         Season == "Spring",
         !(Point_Name %in% c("THG-09", "THG-10", "THG-11", "THG-12", 
                             "THG-13", "THG-14", "THG-15", "THG-16", 
                             "THG-17", "THG-18"))) %>%
  group_by(Species, Point_Name) %>%
  count() %>%
  pivot_wider(names_from = Species, values_from = n)
tnc_site_by_species_spring[is.na(tnc_site_by_species_spring)] = 0
site_species_spring_2021=data.frame(tnc_site_by_species_spring[,-1])

sites_spring_2021 = tnc_site_by_species_spring %>% pull(Point_Name)
transects_spring_2021 = sapply(strsplit(sites_spring_2021,"-"), `[`, 1)
habitat_spring_2021 = as.data.frame(transects_spring_2021) %>%
  mutate(Habitat = case_when(transects_spring_2021 == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             transects_spring_2021 == "CO" ~ "Oak Woodland",
                             transects_spring_2021 == "THG" ~ "Grassland",
                             transects_spring_2021 == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian")) %>%
  dplyr::select(Habitat)
col_spring_2021 = habitat_spring_2021 %>%
  mutate(color = case_when(habitat_spring_2021 == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           habitat_spring_2021 == "Oak Woodland" ~ "grey",
                           habitat_spring_2021 == "Grassland" ~ "green",
                           habitat_spring_2021 == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue")) %>%
    dplyr::select(color)

tnc_spring_2021_shannon = diversity(site_species_spring_2021)

# Ordination using NMDS and random starting points
tnc_spring_2021.mds = metaMDS(site_species_spring_2021, trace = F)
tnc_spring_2021.mds
row.names(tnc_spring_2021.mds$points) = sites_spring_2021

# Some Ordination plots
ordiplot(tnc_spring_2021.mds, type = "n")
ordihull(tnc_spring_2021.mds, groups = habitat_spring_2021$Habitat, draw="polygon",
         col = c("yellow", "green", "brown", "grey", "blue"), label=F)
ordiellipse(tnc_spring_2021.mds, groups = habitat_spring_2021$Habitat, draw="polygon", label=T)
# ordispider(tnc_spring_2021.mds, groups=habitat_spring_2021$Habitat)
orditorp(tnc_spring_2021.mds, display="species", col="red", air=0.01)
orditorp(tnc_2021.mds, display="sites", col = col_2021$color,
   air=0.01,cex=1.25)

# Following my methods from CMP point count talk
tnc_spring_2021.euc = vegdist(site_species_spring_2021, method = "euc")
tnc_spring_2021.ward = hclust(tnc_spring_2021.euc, method = "ward.D")
cor(tnc_spring_2021.euc, cophenetic(tnc_spring_2021.ward))
plot(tnc_spring_2021.ward, labels = sites_spring_2021)

indval_spring_2021 = multipatt(site_species_spring_2021, habitat_spring_2021$Habitat, control = how(nperm = 999))
summary(indval_spring_2021)

## Summer
tnc_site_by_species_summer = tnc_combined_2021 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101,
         Season == "Summer") %>%
  group_by(Species, Point_Name) %>%
  count() %>%
  pivot_wider(names_from = Species, values_from = n)
tnc_site_by_species_summer[is.na(tnc_site_by_species_summer)] = 0
site_species_summer_2021=data.frame(tnc_site_by_species_summer[,-1])

sites_summer_2021 = tnc_site_by_species_summer %>% pull(Point_Name)
transects_summer_2021 = sapply(strsplit(sites_summer_2021,"-"), `[`, 1)
habitat_summer_2021 = as.data.frame(transects_summer_2021) %>%
  mutate(Habitat = case_when(transects_summer_2021 == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             transects_summer_2021 == "CO" ~ "Oak Woodland",
                             transects_summer_2021 == "THG" ~ "Grassland",
                             transects_summer_2021 == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian")) %>%
  dplyr::select(Habitat)
col_summer_2021 = habitat_summer_2021 %>%
  mutate(color = case_when(habitat_summer_2021 == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           habitat_summer_2021 == "Oak Woodland" ~ "grey",
                           habitat_summer_2021 == "Grassland" ~ "green",
                           habitat_summer_2021 == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue")) %>%
    dplyr::select(color)

tnc_summer_2021_shannon = diversity(site_species_summer_2021)

tnc_summer_2021.dis = vegdist(site_species_summer_2021)
range(tnc_summer_2021.dis)
tnc_summer_2021.mds0 = isoMDS(tnc_summer_2021.dis)
stressplot(tnc_summer_2021.mds0, tnc_summer_2021.dis)
# An Ordination plot
p1 = ordiplot(tnc_summer_2021.mds0, type = "points")
row.names(p1$sites) = sites_summer_2021
text(p1, "sites", col = "red")

# Ordination using NMDS and random starting points
tnc_summer_2021.mds = metaMDS(site_species_summer_2021, trace = F)
tnc_summer_2021.mds
row.names(tnc_summer_2021.mds$points) = sites_summer_2021

# Some Ordination plots
ordiplot(tnc_summer_2021.mds, type = "n")
ordihull(tnc_summer_2021.mds, groups = habitat_summer_2021$Habitat, draw="polygon",
         col = c("yellow", "green", "brown", "grey", "blue"), label=F)
ordiellipse(tnc_summer_2021.mds, groups = habitat_summer_2021$Habitat, draw="polygon", label=T)
# ordispider(tnc_summer_2021.mds, groups=habitat_summer_2021$Habitat)
orditorp(tnc_summer_2021.mds, display="species", col="red", air=0.01)
orditorp(tnc_2021.mds, display="sites", col = col_2021$color,
   air=0.01,cex=1.25)

# Following my methods from CMP point count talk
tnc_summer_2021.euc = vegdist(site_species_summer_2021, method = "euc")
tnc_summer_2021.ward = hclust(tnc_summer_2021.euc, method = "ward.D")
cor(tnc_summer_2021.dis, cophenetic(tnc_summer_2021.ward))
plot(tnc_summer_2021.ward, labels = sites_summer_2021)

indval_summer_2021 = multipatt(site_species_summer_2021, habitat_summer_2021$Habitat, control = how(nperm = 999))
summary(indval_summer_2021)

## Fall
tnc_site_by_species_fall = tnc_combined_2021 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101,
         Season == "Fall") %>%
  group_by(Species, Point_Name) %>%
  count() %>%
  pivot_wider(names_from = Species, values_from = n)
tnc_site_by_species_fall[is.na(tnc_site_by_species_fall)] = 0
site_species_summer_2021=data.frame(tnc_site_by_species_fall[,-1])


## Rarefaction
abundance_all_2021 = rowSums(site_species)
names(abundance_all_2021)
raremin_all_2021 = min(abundance_all_2021)
sRare15_2021 = rarefy(site_species, 15)
rownames(site_species)= sites_2021
rarecurve(site_species, col = col_2021$color, label = T)
rare_tidy = rarecurve(site_species, tidy = T)

rare_tidy$Transect = sapply(strsplit(as.character(rare_tidy$Site),"-"), `[`, 1)
rare_tidy = rare_tidy %>%
  mutate(Habitat = case_when(Transect == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             Transect == "CO" ~ "Oak Woodland",
                             Transect == "THG" ~ "Grassland",
                             Transect == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian"))

rare_tidy = rare_tidy %>%
  mutate(Color = case_when(Habitat == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           Habitat == "Oak Woodland" ~ "grey",
                           Habitat == "Grassland" ~ "green",
                           Habitat == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue"))

ggplot(data = rare_tidy, aes(Sample, Species,
                             # color = Habitat,
                             group = Site)) +
  geom_path(size = 1, show.legend = F) +
  labs(title = "Species accumulation curves",
       # subtitle = "(each line is a point count site)",
       x = "Sample Size") +
  facet_wrap(~Habitat, nrow = 2)


# Summer

rownames(site_species_summer_2021)= sites_summer_2021
rare_tidy_summer_2021 = rarecurve(site_species_summer_2021, tidy = T)
rare_tidy_summer_2021$Transect = sapply(strsplit(as.character(rare_tidy_summer_2021$Site),"-"), `[`, 1)
rare_tidy_summer_2021 = rare_tidy_summer_2021 %>%
  mutate(Habitat = case_when(Transect == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             Transect == "CO" ~ "Oak Woodland",
                             Transect == "THG" ~ "Grassland",
                             Transect == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian"))

rare_tidy_summer_2021 = rare_tidy_summer_2021 %>%
  mutate(Color = case_when(Habitat == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           Habitat == "Oak Woodland" ~ "grey",
                           Habitat == "Grassland" ~ "green",
                           Habitat == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue"))

ggplot(data = rare_tidy_summer_2021, aes(Sample, Species,
                             # color = Habitat,
                             group = Site)) +
  geom_path(size = 1, show.legend = F) +
  labs(title = "Species accumulation curves",
       # subtitle = "(each line is a point count site)",
       x = "Sample Size") +
  facet_wrap(~Habitat, nrow = 2)


```

### 2021 - 2023

Most changes were done by Sasha in spreadsheets.

This year included summary of 3 years.

Columns 31-40 appear to have been used to sort, proof and combine years of data and are selected out for now.

Todo:

* leave empid as UEFL


```{r data23}

# Read in Point count data
TNC_Point_Count_Data_2021_2023 <- read_excel("C:/Users/PatrickLorch/SSRS/Southern Sierra Research Station - Documents/Projects/TNC Bird Surveys - Caliente_WB/Data/2023/3 years of data/TNC Point Count Data 2021-2023.xlsx",
sheet = "2021-2023", col_types = c("numeric",
"text", "text", "text", "text", "date",
"numeric", "text", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "text", "text", "text",
"text", "text", "text", "text", "numeric",
"numeric", "text", "text", "text",
"text", "numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric"))

names(TNC_Point_Count_Data_2021_2023)

# Removes columns 31-40
tnc_pc_2021_2023 = TNC_Point_Count_Data_2021_2023 %>%
  select(1:30)

# Sanity check
table(tnc_pc_2021_2023$Species, tnc_pc_2021_2023$Location_Code)

# Set dummy variable for No birds at a point (NOBI)
tnc_combined_2021_2023 = tnc_pc_2021_2023 %>%
  mutate(nobi = ifelse(Species == "NOBI", 1, 0))

# Mark unknowns
## Change older Unknown Empidonax to new code
which(tnc_combined_2021_2023$Species == "UEFL")
which(tnc_combined_2021_2023$Species == "UNEM")
tnc_combined_2021_2023 = tnc_combined_2021_2023 %>%
  mutate(Species = ifelse(Species == "UEFL", "UNEM", Species))

tnc_combined_2021_2023 = tnc_combined_2021_2023 %>%
  mutate(known = substr(Species,1,2) != "UN")

# Habitat type variable was added manually?
table(tnc_combined_2021_2023$Habitat, tnc_combined_2021_2023$Location_Code)

# Season type variable was added manually?
# In other years we created a Season column from Visit
# This code converted anything that was not a 1, 2, or 3 to Fall
# So this correctly dealt with F1, F2 visits, for example
table(tnc_combined_2021_2023$Season, tnc_combined_2021_2023$Visit)

tnc_combined_2021_2023 = tnc_combined_2021_2023 %>%
  mutate(point_vs_88 = case_when(Point_Name == "88" ~ 2,
                                 TRUE ~ 1)
         )

# Replace blank Number with 1
which(is.na(tnc_combined_2021_2023$Number))
tnc_combined_2021_2023 = tnc_combined_2021_2023 %>%
  mutate(Number = ifelse(is.na(Number), 1, Number))

# We need to look at some stuff by year, so adding this here
tnc_combined_2021_2023 = tnc_combined_2021_2023 %>%
  mutate(year = year(tnc_combined_2021_2023$Date))

# To skip the above steps, we can import this file.
write.csv(tnc_combined_2021_2023, "tnc_combined_2021_2023.csv")           

# Checking that THG-19 and THG-9 show up in correct years
pt_name_by_year_21_23 = tnc_combined_2021_2023 %>%
  filter(known == T, nobi == 0) %>%
  group_by(Point_Name, year) %>%
  count() %>%
  spread(Point_Name, n)


# summaries 
sp_by_hab_21_23 = tnc_combined_2021_2023 %>%
  filter(known == T, nobi == 0) %>%
  group_by(Species, Habitat) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_21_23, "sp_by_hab_21_23.csv")

sp_counts_by_hab_21_23 = colSums(sp_by_hab_21_23 != 0, na.rm = T)
write.csv(as.data.frame(sp_counts_by_hab_21_23), "sp_counts_by_hab_21_23.csv")

# Yearly
sp_by_hab_year_21_23 = tnc_combined_2021_2023 %>%
  filter(known == T, nobi == 0) %>%
  group_by(Species, Habitat, year) %>%
  count() %>%
  pivot_wider(names_from = c(Habitat, year), values_from = n)
write.csv(sp_by_hab_year_21_23, "3 years/sp_by_hab_year_21_23.csv")

sp_counts_by_hab_year_21_23 = colSums(sp_by_hab_year_21_23 != 0, na.rm = T)
write.csv(as.data.frame(sp_counts_by_hab_year_21_23), "3 years/sp_counts_by_hab_year_21_23.csv")

sp_by_hab_season_21_23 = tnc_combined_2021_2023 %>%
  filter(known == T, nobi == 0) %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_season_21_23, "sp_by_hab_season_21_23.csv")

sp_by_hab_season_wo88_21_23 = tnc_combined_2021_2023 %>%
  filter(known == T, nobi == 0, point_vs_88 == 1) %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_season_wo88_21_23, "sp_by_hab_season_wo88_21_23.csv")

caliente_observer_21_23 = tnc_combined_2021_2023 %>%
  select("Date", "Location_Code", "Observer") %>%
  unique() %>%
  left_join(caliente_location_codes, by = c("Location_Code" = "site_code")) %>%
  left_join(surveyor, by = c("Observer" = "initials")) %>%
  select(Date = Date, Site = site_name, Surveyor = name)
write.csv(caliente_observer_21_23, "caliente_observer_21_23.csv")

bird_sp = read.csv("birds_202210041059.csv")
bird_sp = bird_sp %>%
  select(spec, commonname, sciname)

sp_list_points_21_23 = tnc_combined_2021_2023 %>%
  filter(known == T, nobi == 0, point_vs_88 == 1) %>%
  select(Species, Season)%>%
  distinct() %>%
  arrange(Season, Species) %>%
  left_join(bird_sp, by = c("Species" = "spec"))

sp_list_points_21_23_by_season = sp_list_points_21_23 %>%
  mutate(count = 1, nobi = 0) %>%
  pivot_wider(names_from = Season, values_from = count)
# generate totals by season if time permits
write.csv(sp_list_points_21_23_by_season, "sp_list_points_21_23_by_season.csv")
colSums(sp_list_points_21_23_by_season[,5:7], na.rm = T)

sp_list_88_21_23 = tnc_combined_2021_2023 %>%
  filter(known == T, nobi == 0, point_vs_88 == 2) %>%
  select(Species, Season)%>%
  distinct() %>%
  arrange(Season, Species) %>%
  left_join(bird_sp, by = c("Species" = "spec"))

sp_list_88_21_23_by_season = sp_list_88_21_23 %>%
  mutate(count = 1) %>%
  pivot_wider(names_from = Season, values_from = count)
write.csv(sp_list_88_21_23_by_season, "sp_list_88_21_23_by_season.csv")
colSums(sp_list_88_21_23_by_season[,4:6], na.rm = T)

# Anti join to find unique species in 88s
sp_list_88_21_23_unique = sp_list_88_21_23 %>%
  anti_join(sp_list_points_21_23, join_by("Species", "Season")) %>%
  mutate(count = 1) %>%
  pivot_wider(names_from = Season, values_from = count)
write.csv(sp_list_88_21_23_unique, "sp_list_88_21_23_unique.csv")
colSums(sp_list_88_21_23_unique[,4:6], na.rm = T)

```

### Filtering by distance and flyovers

Further truncating the data by filtering out observations greater than 100 meters, and sometimes 88 and flyovers.

```{r distancecutoff}
sp_by_hab_100mflyover_21_23 = tnc_combined_2021_2023 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101) %>%
  group_by(Species, Habitat) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_100mflyover_21_23, "DistanceCutoff/sp_by_hab_100mflyover_21_23.csv")

sp_counts_by_hab_100mflyover_21_23 = colSums(sp_by_hab_100mflyover_21_23 != 0, na.rm = T)
write.csv(as.data.frame(sp_counts_by_hab_100mflyover_21_23), "DistanceCutoff/sp_counts_by_hab_100mflyover_21_23.csv")

sp_by_hab_100mflyover_season_21_23 = tnc_combined_2021_2023 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101) %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_100mflyover_season_21_23, "DistanceCutoff/sp_by_hab_100mflyover_season_21_23.csv")

sp_by_hab_100mflyover_season_wo88_21_23 = tnc_combined_2021_2023 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101, point_vs_88 == 1) %>%
  group_by(Species, Habitat, Season) %>%
  # distinct(Species) %>%
  count() %>%
  group_by(Habitat, Season) %>%
  count() %>%
  spread(Habitat, n)
write.csv(sp_by_hab_100mflyover_season_wo88_21_23, "DistanceCutoff/sp_by_hab_100mflyover_season_wo88_21_23.csv")

```


#### Community analyses 2021-2023

We want to compare habitat types in a dendrogram.

Following the vegan community vignette:
https://www.mooreecology.com/uploads/2/4/2/1/24213970/vegantutor.pdf

```{r community2123}
#install.packages("vegan")
library(MASS)
library(vegan)
library(tidyr)
library(dplyr)
library(plotrix)
library(indicspecies)
library(ggplot2)

tnc_site_by_species_all_21_23 = tnc_combined_2021_2023 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101) %>%
  group_by(Species, Point_Name) %>%
  count() %>%
  pivot_wider(names_from = Species, values_from = n)

tnc_site_by_species_all_21_23[is.na(tnc_site_by_species_all_21_23)] = 0
site_species_all_21_23=data.frame(tnc_site_by_species_all_21_23[,-1])

tnc_site_by_species_21_23 = tnc_site_by_species_all_21_23 %>%
  filter(!(Point_Name %in% c("THG-09",
                                        "THG-10",
                                        "THG-11",
                                        "THG-12",
                                        "THG-13",
                                        "THG-14",
                                        "THG-15",
                                        "THG-16",
                                        "THG-17",
                                        "THG-18")))

tnc_site_by_species_21_23[is.na(tnc_site_by_species_21_23)] = 0
site_species_21_23=data.frame(tnc_site_by_species_21_23[,-1])

sites_21_23 = tnc_site_by_species_21_23 %>% pull(Point_Name)
transects_21_23 = sapply(strsplit(sites_21_23,"-"), `[`, 1)
habitat_21_23 = as.data.frame(transects_21_23) %>%
  mutate(Habitat = case_when(transects_21_23 == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             transects_21_23 == "CO" ~ "Oak Woodland",
                             transects_21_23 == "THG" ~ "Grassland",
                             transects_21_23 == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian")) %>%
  dplyr::select(Habitat)
col_21_23 = habitat_21_23 %>%
  mutate(color = case_when(habitat_21_23 == "Buckwheat/Goldenbush Scrub" ~ "orange",
                           habitat_21_23 == "Oak Woodland" ~ "grey",
                           habitat_21_23 == "Grassland" ~ "yellow",
                           habitat_21_23 == "Oak Savannah" ~ "blue",
                           TRUE ~ "green")) %>%
    dplyr::select(color)

tnc_21_23_shannon = diversity(site_species_21_23)
tnc_21_23_simpson = diversity(site_species_21_23, index = "simpson")
tnc_21_23_diversity_df = data.frame(points_sim = names(tnc_21_23_simpson),
                                  tnc_21_23_simpson,
                                  points_shan = names(tnc_21_23_shannon),
                                  tnc_21_23_shannon)
transects_div_21_23 = sapply(strsplit(tnc_21_23_diversity_df$points_sim,"-"), `[`, 1)
habitat_div_21_23 = as.data.frame(transects_div_21_23) %>%
  mutate(Habitat = case_when(transects_div_21_23 == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             transects_div_21_23 == "CO" ~ "Oak Woodland",
                             transects_div_21_23 == "THG" ~ "Grassland",
                             transects_div_21_23 == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian")) %>%
  dplyr::select(Habitat)
tnc_21_23_diversity_df = data.frame(tnc_21_23_diversity_df, Habitat = habitat_div_21_23)

# Simpson
tnc_21_23_diversity_simpson.aov = aov(formula = log(tnc_21_23_simpson + 0.001) ~ Habitat, data = tnc_21_23_diversity_df)

sink("tnc_21_23_diversity_simpson.txt")
summary(tnc_21_23_diversity_simpson.aov)
TukeyHSD(tnc_21_23_diversity_simpson.aov)
sink()

boxplot(formula = log(tnc_21_23_simpson + 0.001) ~ Habitat, data = tnc_21_23_diversity_df)

# Shannon
tnc_21_23_diversity_shannon.aov = aov(formula = tnc_21_23_shannon ~ Habitat, data = tnc_21_23_diversity_df)

sink("tnc_21_23_diversity_shannon.txt")
summary(tnc_21_23_diversity_shannon.aov)
TukeyHSD(tnc_21_23_diversity_shannon.aov)
sink()

boxplot(formula = tnc_21_23_shannon ~ Habitat, data = tnc_21_23_diversity_df)


# Calculate relative abundance by site
tnc_21_23_relabund = decostand(site_species_21_23, method = "total")
apply(tnc_21_23_relabund, 1, sum)


tnc_21_23.dis = vegdist(site_species_21_23)
tnc_21_23.mds0 = isoMDS(tnc_21_23.dis)
stressplot(tnc_21_23.mds0, tnc_21_23.dis)
# An Ordination plot
p1 = ordiplot(tnc_21_23.mds0, type = "points")
row.names(p1$sites) = sites_21_23
text(p1, "sites", col = "red")

# Ordination using NMDS and random starting points
tnc_21_23.mds = metaMDS(site_species_21_23, trace = F)
tnc_21_23.mds
row.names(tnc_21_23.mds$points) = sites_21_23
p1 = plot(tnc_21_23.mds, disp = "sp")
# This should reduce business of species labeling, but it takes forever with this ordiplot
# identify(p1, "sp")
hab_fit_2021 = envfit(tnc_21_23.mds, habitat_21_23, permu = 999)
p1 = plot(tnc_21_23.mds, disp = "sp")
plot(hab_fit_2021, p.max = 0.1)

# Some Ordination plots
orplt = ordiplot(tnc_21_23.mds, type = "n", 
                 ylim = c(-1, 1), xlim = c(-2, 2.5))
ordiellipse_tnc2123 = ordiellipse(tnc_21_23.mds, groups = habitat_21_23$Habitat, draw="polygon",
         col = c("orange", "yellow", "blue", "grey", "green"), label =F) # , cex = 0.5
summary(ordiellipse_tnc2123)
legend("bottomright", 
       legend = unique(habitat_21_23$Habitat), 
       col = c("green", "orange", "grey", "yellow", "blue"),
       pch = 15, cex = 1)

ordihull(tnc_21_23.mds, groups = habitat_21_23$Habitat, draw="polygon",
col = c("yellow", "green", "brown", "grey", "blue"), label=F)

# ordispider(tnc_21_23.mds, groups=habitat_21_23$Habitat)
# write
orditorp(tnc_21_23.mds, display="species", col="red", air=0.01)
orditorp(tnc_21_23.mds, display="sites", col = col_21_23$color,
   air=0.01,cex=1.25)

orplt2 = ordiplot(tnc_21_23.mds, type = "n", 
                 ylim = c(-2, 2), xlim = c(-3, 3))
points(tnc_21_23.mds, 
       col = c("yellow", "green", "brown", "grey", "blue"), 
       pch = as.numeric(as.factor(habitat_21_23$Habitat)))
ordihull(tnc_21_23.mds, groups = habitat_21_23$Habitat, draw="polygon",
col = c("yellow", "green", "brown", "grey", "blue"), label=F, alpha = 127)
legend("bottomright", 
       legend = unique(habitat_21_23$Habitat), 
       col = c("green", "orange", "grey", "yellow", "blue"),
       pch = 15, cex = 1)


tnc_21_23.clus = hclust(tnc_21_23.dis, "single")
plot(tnc_21_23.clus, labels = sites_21_23)

tnc_21_23.cluc = hclust(tnc_21_23.dis, "complet")
plot(tnc_21_23.cluc, labels = sites_21_23)

tnc_21_23.clua = hclust(tnc_21_23.dis, "average")
plot(tnc_21_23.clua, labels = sites_21_23)

range(tnc_21_23.dis)

cor(tnc_21_23.dis, cophenetic(tnc_21_23.clus))
cor(tnc_21_23.dis, cophenetic(tnc_21_23.cluc))
cor(tnc_21_23.dis, cophenetic(tnc_21_23.clua))

# Following my methods from CMP point count talk
tnc_21_23.euc = vegdist(site_species_21_23, method = "euc")
tnc_21_23.ward = hclust(tnc_21_23.euc, method = "ward.D")
plot(tnc_21_23.ward, labels = sites_21_23)
cor(tnc_21_23.euc, cophenetic(tnc_21_23.ward))

indval_21_23 = multipatt(site_species_21_23, habitat_21_23$Habitat, control = how(nperm = 999))
sink(file = "indval_21_23_sum.txt")
summary(indval_21_23) 
sink()

## By season
# Needed to eliminate THG-09 - 18 for spring 2021 and possibly 2022

## Spring
tnc_site_by_species_21_23_spring = tnc_combined_2021_2023 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101,
         Season == "Spring",
         !(Point_Name %in% c("THG-09", "THG-10", "THG-11", "THG-12", 
                             "THG-13", "THG-14", "THG-15", "THG-16", 
                             "THG-17", "THG-18"))) %>%
  group_by(Species, Point_Name) %>%
  count() %>%
  pivot_wider(names_from = Species, values_from = n)
tnc_site_by_species_21_23_spring[is.na(tnc_site_by_species_21_23_spring)] = 0
site_species_spring_21_23=data.frame(tnc_site_by_species_21_23_spring[,-1])

sites_spring_21_23 = tnc_site_by_species_21_23_spring %>% pull(Point_Name)
transects_spring_21_23 = sapply(strsplit(sites_spring_21_23,"-"), `[`, 1)
habitat_spring_21_23 = as.data.frame(transects_spring_21_23) %>%
  mutate(Habitat = case_when(transects_spring_21_23 == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             transects_spring_21_23 == "CO" ~ "Oak Woodland",
                             transects_spring_21_23 == "THG" ~ "Grassland",
                             transects_spring_21_23 == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian")) %>%
  dplyr::select(Habitat)
col_spring_21_23 = habitat_spring_21_23 %>%
  mutate(color = case_when(habitat_spring_21_23 == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           habitat_spring_21_23 == "Oak Woodland" ~ "grey",
                           habitat_spring_21_23 == "Grassland" ~ "green",
                           habitat_spring_21_23 == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue")) %>%
    dplyr::select(color)

tnc_spring_21_23_shannon = diversity(site_species_spring_21_23)

# Ordination using NMDS and random starting points
tnc_spring_21_23.mds = metaMDS(site_species_spring_21_23, trace = F)
tnc_spring_21_23.mds
row.names(tnc_spring_21_23.mds$points) = sites_spring_21_23

# Some Ordination plots
ordiplot(tnc_spring_21_23.mds, type = "n")
ordihull(tnc_spring_21_23.mds, groups = habitat_spring_21_23$Habitat, draw="polygon",
         col = c("yellow", "green", "brown", "grey", "blue"), label=F)
ordiellipse(tnc_spring_21_23.mds, groups = habitat_spring_21_23$Habitat, draw="polygon", label=T)
# ordispider(tnc_spring_21_23.mds, groups=habitat_spring_21_23$Habitat)
orditorp(tnc_spring_21_23.mds, display="species", col="red", air=0.01)
orditorp(tnc_spring_21_23.mds, display="sites", col = col_21_23$color,
   air=0.01,cex=1.25)

# Following my methods from CMP point count talk
tnc_spring_21_23.euc = vegdist(site_species_spring_21_23, method = "euc")
tnc_spring_21_23.ward = hclust(tnc_spring_21_23.euc, method = "ward.D")
cor(tnc_spring_21_23.euc, cophenetic(tnc_spring_21_23.ward))
plot(tnc_spring_21_23.ward, labels = sites_spring_21_23)

indval_spring_21_23 = multipatt(site_species_spring_21_23, habitat_spring_21_23$Habitat, control = how(nperm = 999))
summary(indval_spring_21_23)

## Summer
tnc_site_by_species_21_23_summer = tnc_combined_2021_2023 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101,
         Season == "Summer") %>%
  group_by(Species, Point_Name) %>%
  count() %>%
  pivot_wider(names_from = Species, values_from = n)
tnc_site_by_species_21_23_summer[is.na(tnc_site_by_species_21_23_summer)] = 0
site_species_summer_21_23=data.frame(tnc_site_by_species_21_23_summer[,-1])

sites_summer_21_23 = tnc_site_by_species_21_23_summer %>% pull(Point_Name)
transects_summer_21_23 = sapply(strsplit(sites_summer_21_23,"-"), `[`, 1)
habitat_summer_21_23 = as.data.frame(transects_summer_21_23) %>%
  mutate(Habitat = case_when(transects_summer_21_23 == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             transects_summer_21_23 == "CO" ~ "Oak Woodland",
                             transects_summer_21_23 == "THG" ~ "Grassland",
                             transects_summer_21_23 == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian")) %>%
  dplyr::select(Habitat)
col_summer_21_23 = habitat_summer_21_23 %>%
  mutate(color = case_when(habitat_summer_21_23 == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           habitat_summer_21_23 == "Oak Woodland" ~ "grey",
                           habitat_summer_21_23 == "Grassland" ~ "green",
                           habitat_summer_21_23 == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue")) %>%
    dplyr::select(color)

tnc_summer_21_23_shannon = diversity(site_species_summer_21_23)

tnc_summer_21_23.dis = vegdist(site_species_summer_21_23)
range(tnc_summer_21_23.dis)
# This fails
tnc_summer_21_23.mds0 = isoMDS(tnc_summer_21_23.dis)
stressplot(tnc_summer_21_23.mds0, tnc_summer_21_23.dis)
# An Ordination plot
p1 = ordiplot(tnc_summer_21_23.mds0, type = "points")
row.names(p1$sites) = sites_summer_21_23
text(p1, "sites", col = "red")

# Ordination using NMDS and random starting points
tnc_summer_21_23.mds = metaMDS(site_species_summer_21_23, trace = F)
tnc_summer_21_23.mds
row.names(tnc_summer_21_23.mds$points) = sites_summer_21_23

# Some Ordination plots
ordiplot(tnc_summer_21_23.mds, type = "n")
ordihull(tnc_summer_21_23.mds, groups = habitat_summer_21_23$Habitat, draw="polygon",
         col = c("yellow", "green", "brown", "grey", "blue"), label=F)
ordiellipse(tnc_summer_21_23.mds, groups = habitat_summer_21_23$Habitat, draw="polygon", label=T)
# ordispider(tnc_summer_21_23.mds, groups=habitat_summer_21_23$Habitat)
orditorp(tnc_summer_21_23.mds, display="species", col="red", air=0.01)
orditorp(tnc_summer_21_23.mds, display="sites", col = col_21_23$color,
   air=0.01,cex=1.25)

# Following my methods from CMP point count talk
tnc_summer_21_23.euc = vegdist(site_species_summer_21_23, method = "euc")
tnc_summer_21_23.ward = hclust(tnc_summer_21_23.euc, method = "ward.D")
cor(tnc_summer_21_23.dis, cophenetic(tnc_summer_21_23.ward))
plot(tnc_summer_21_23.ward, labels = sites_summer_21_23)

indval_summer_21_23 = multipatt(site_species_summer_21_23, habitat_summer_21_23$Habitat, control = how(nperm = 999))
summary(indval_summer_21_23)

## Fall
tnc_site_by_species_21_23_fall = tnc_combined_2021_2023 %>%
  filter(is.na(Fly_Over), known == T, nobi == 0, Distance < 101,
         Season == "Fall") %>%
  group_by(Species, Point_Name) %>%
  count() %>%
  pivot_wider(names_from = Species, values_from = n)
tnc_site_by_species_21_23_fall[is.na(tnc_site_by_species_21_23_fall)] = 0
site_species_summer_21_23=data.frame(tnc_site_by_species_21_23_fall[,-1])


## Rarefaction
abundance_all_21_23 = rowSums(site_species_21_23)
names(abundance_all_21_23)
raremin_all_21_23 = min(abundance_all_21_23)
sRare15_21_23 = rarefy(site_species_21_23, 15)
rownames(site_species_21_23)= sites_21_23
rarecurve(site_species_21_23, col = col_21_23$color, label = T)
rare_tidy_21_23 = rarecurve(site_species_21_23, tidy = T)

rare_tidy_21_23$Transect = sapply(strsplit(as.character(rare_tidy_21_23$Site),"-"), `[`, 1)
rare_tidy_21_23 = rare_tidy_21_23 %>%
  mutate(Habitat = case_when(Transect == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             Transect == "CO" ~ "Oak Woodland",
                             Transect == "THG" ~ "Grassland",
                             Transect == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian"))

rare_tidy_21_23 = rare_tidy_21_23 %>%
  mutate(Color = case_when(Habitat == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           Habitat == "Oak Woodland" ~ "grey",
                           Habitat == "Grassland" ~ "green",
                           Habitat == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue"))

ggplot(data = rare_tidy_21_23, aes(Sample, Species,
                             # color = Habitat,
                             group = Site)) +
  geom_path(size = 1, show.legend = F) +
  labs(title = "Species accumulation curves",
       # subtitle = "(each line is a point count site)",
       x = "Sample Size") +
  facet_wrap(~Habitat, nrow = 2)

## STOPPED here 2/23/24
## This did not work when the number of rownames was way less than sites
# Summer

rownames(site_species_summer_21_23)= sites_summer_21_23
rare_tidy_summer_21_23 = rarecurve(site_species_summer_21_23, tidy = T)
rare_tidy_summer_21_23$Transect = sapply(strsplit(as.character(rare_tidy_summer_21_23$Site),"-"), `[`, 1)
rare_tidy_summer_21_23 = rare_tidy_summer_21_23 %>%
  mutate(Habitat = case_when(Transect == "CBG" ~ "Buckwheat/Goldenbush Scrub",
                             Transect == "CO" ~ "Oak Woodland",
                             Transect == "THG" ~ "Grassland",
                             Transect == "THO" ~ "Oak Savannah",
                             TRUE ~ "Riparian"))

rare_tidy_summer_21_23 = rare_tidy_summer_21_23 %>%
  mutate(Color = case_when(Habitat == "Buckwheat/Goldenbush Scrub" ~ "yellow",
                           Habitat == "Oak Woodland" ~ "grey",
                           Habitat == "Grassland" ~ "green",
                           Habitat == "Oak Savannah" ~ "brown",
                           TRUE ~ "blue"))

ggplot(data = rare_tidy_summer_21_23, aes(Sample, Species,
                             # color = Habitat,
                             group = Site)) +
  geom_path(size = 1, show.legend = F) +
  labs(title = "Species accumulation curves",
       # subtitle = "(each line is a point count site)",
       x = "Sample Size") +
  facet_wrap(~Habitat, nrow = 2)


```

